---
title: "DS-Capstone-Milestone1"
author: "Pat O'Keeffe"
date: "25/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corpus)
library(caret)
library(gsubfn)
library(stopwords)
setwd('~/data-science-capstone/')
```

## This is the first milestone for the data science capstone project.

The goal of this milestone report is to do the following:
1. Demonstrate download of the data and load into data structures within R.
2. Provide some summary statistics about the data that is downloaded.
3. Report some interesting findings so far.
4. Describe plans for creating a prediction algorithm and a Shiny app.

### Download of the data.

The data was downloaded from the swiftkey in a zip file and extracted to a data directory within the working directory. 
The following code is used to load the data of each file into a dataframe.

```{r data-load}
twitter <- readLines('data/en_US/en_US.twitter.txt')
news <- readLines('data/en_US/en_US.news.txt')
blogs <- readLines('data/en_US/en_US.blogs.txt')
blogs_cp <- scan(file='data/en_US/en_US.blogs.txt',sep='\n',what='char',quote='')
```

### Summary Statistics about the data downloaded.

Here we will do a frequency distribution of the words in the twitter feed. I also created a stop list of words that should be excluded from the frequency. Also remove words that only occur once and finally provide a plot distribution of the top 100 words.

```{r frequency, echo=FALSE}
blogs_cp <- tolower(blogs_cp)
words.list <- strsplit(blogs_cp,'\\W+')
words.vector <- unlist(words.list)

#Create a stop list of uninteresting words
stop.list <- c('the','a','of','and','is','be','are')
words.vector2 <- words.vector[!words.vector%in%stop.list]
freq.list <- table(words.vector2)
sorted.freq.list <- sort(freq.list,decreasing = T)

#remove hapax legomena (words that occur only once in the corpora)
sorted.freq.list <- sorted.freq.list[sorted.freq.list>1] 

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
